---
title: "Chapter 5: Conditional Standard Errors and Classification Accuracy & Consistency"
author: "BRT"
date: "8/20/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library("readxl")
library(cacIRT)
library(kableExtra)
library(janitor)
library(knitr)
library(here)

# This chapter requires the following data files:
# datastore/2020PASA_ScalingAnalysis/par-estimates-2020/person-estimates
# datastore/2020PASA_ScalingAnalysis/par-estimates-2020/item-estimates
# datastore/2020PASA_ScalingAnalysis/cut-scores/plc-cuts.xlsx
# datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_ELA_Final_2020.csv
# datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Math_Final_2020.csv
# datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Science_Final_2020.csv

p_files <- fs::dir_ls(here::here("data", "person-estimates"))

p <- map_df(p_files, read_csv, .id = "file") %>%
  mutate(file = gsub(here::here("data", "person-estimates"), "", file),
         grade = as.numeric(gsub(".+g(\\d?\\d).+", "\\1", file)),
         tier = as.numeric(gsub(".+\\dt(\\d).+", "\\1", file)),
         content = gsub(".+-(.+)\\..+", "\\1", file)) %>%
  select(student_id, grade, tier, content, TotalScore:SEM)

cuts <- read_excel(here::here("data", "cut", "plc-cuts.xlsx"))

ela <- read_csv(here::here("data", "score", "Scoring_ELA_Final_2020.csv")) %>%
  clean_names()

math <- read_csv(here::here("data", "score", "Scoring_Math_Final_2020.csv")) %>%
  clean_names()

science <- read_csv(here::here("data", "score", "Scoring_Science_Final_2020.csv")) %>%
  clean_names()
```

# 7.1 Conditional Standard Errors 

The average standard error of measurement (SEM) associated with each proficiency
level cut score for 2019-20 student data are presented in the tables below. 

```{r stand_error_measure, results="asis"}

cutoff <- p %>%
  mutate(ScaleScore = round(ScaleScore)) %>%
  group_by(tier, grade, content) %>%
  nest() %>%
  left_join(cuts)

compute_sem <- function(model, c1, c2, c3) {
  l <- list(c1, c2, c3)
  sems <- map_dbl(l, ~predict(model, newdata = data.frame(ScaleScore = .x)))
  data.frame(cut1_se = sems[1],
             cut2_se = sems[2],
             cut3_se = sems[3])
}

cut_score <- cutoff %>%
  mutate(m = map(data, ~lm(SEM ~ poly(ScaleScore, 7), na.omit(.x))),
         sem = pmap(list(m, cut1, cut2, cut3), compute_sem))

# Present as table (round SE)
se_table <- cut_score %>%
  select(grade:content, starts_with("cut"), sem) %>%
  unnest(sem) %>%
  rename(cut1_cut = cut1,
         cut2_cut = cut2,
         cut3_cut = cut3) %>%
  pivot_longer(starts_with("cut"),
               names_to = c("cutt", "var"),
               names_sep = "_",
               values_to = "val") %>%
  mutate(
    cutt = recode(cutt,
                  cut1 = "Novice - Emerging",
                  cut2 = "Emerging - Proficient",
                  cut3 = "Proficient - Advanced"),
    cutt = fct_relevel(cutt, "Novice - Emerging", "Emerging - Proficient"),
    content = recode(content,
                     ela = "ELA",
                     math = "Math",
                     science = "Science")) %>% 
  pivot_wider(names_from = var, values_from = val) %>%
  select(content, grade, tier, cutt, cut, se) %>% 
  mutate(se = round(se, 0)) %>% 
  rename(Grade = grade,
         Tier = tier,
         `Proficency Level Cut`   = cutt,
         `Cut Score` = cut,
         SE = se) %>% 
  arrange(content, Grade, Tier, `Proficency Level Cut`) %>% 
  split(.$content) %>%
  map(~ungroup(.) %>% select(-content))

walk2(se_table,
      paste0(c("ELA", "Math", "Science"),
             " Conditional Standard Errors"),
      ~knitr::kable(.x, 
                    caption = .y) %>%
        kable_styling(full_width = TRUE,
                      latex_options = "hold_position") %>%
        print())

```

# 7.2 Accuracy & Consistency

```{r classification_accuracy}
# p_files <- fs::dir_ls(here::here("data", "item-estimates"))

# theta and SE from p_files

i_files <- fs::dir_ls(here::here("data", "item-estimates"))

 
i <- map_df(i_files, read_csv, .id = "file") %>%
  mutate(file = gsub(here::here("data", "item-estimates"), "", file),
         grade = as.numeric(gsub(".+g(\\d?\\d).+", "\\1", file)),
         tier = as.numeric(gsub(".+\\dt(\\d).+", "\\1", file)),
         content = gsub(".+-(.+)\\..+", "\\1", file)) %>%
  select(item, grade, tier, content, a_estimate, b_estimate, g_estimate)

cuts_theta <- ela %>% 
  select(student_id, grade, tier, ela = plc) %>% 
  left_join(select(math, student_id, math = plc)) %>% 
  left_join(select(science, student_id, science = plc)) %>% 
  pivot_longer(
    cols = c(ela, math, science),
    names_to = "content",
    values_to = "amo_lvl"
  ) %>% 
  left_join(select(p, student_id, grade, tier, content, Theta)) %>% 
  group_by(grade, tier, content, amo_lvl) %>% 
  summarize(cutscore = min(Theta, na.rm = TRUE)) %>% 
  filter(amo_lvl != "Novice") 

cuts_i <- i %>%
  select(content, grade, a_estimate, b_estimate, g_estimate) %>% 
  nest(-content, -grade) %>%
  mutate(ip = map(data, 
                  ~as.matrix(data.frame(.x$a_estimate, .x$b_estimate, .x$g_estimate))),
         content = tolower(content)) %>%
  left_join(cuts_theta)

abils <- p %>%
  drop_na(Theta) %>% #added this line to remove NA to compute marginal below
  nest(-content, -grade) %>%
  rename(persons = data)

cuts_join <- left_join(cuts_i, abils)


acc_con <- cuts_join %>%
  filter(!is.infinite(cutscore)) %>% 
  mutate(theta = map(persons, "Theta"),
         se    = map(persons, "SE"),
         marginal = pmap(list(cutscore, ip, theta, se),
                         function(cutscore, ip, theta, se)
                           class.Rud(cutscore, ip, theta, se, D = 1)$Marginal),
         accuracy = map_dbl(marginal, 1),
         consistency = map_dbl(marginal, 2)) %>% 
  select(content, grade, tier, amo_lvl, accuracy, consistency) %>%
  mutate(content = recode(content,
                     ela = "ELA",
                     math = "Math",
                     science = "Science"),
         amo_lvl = recode(amo_lvl,
                  Emerging = "Novice - Emerging",
                  Proficient = "Emerging - Proficient",
                  Advanced = "Proficient - Advanced"),
         amo_lvl = fct_relevel(amo_lvl, "Novice - Emerging", "Emerging - Proficient")) %>% 
  rename(Grade = grade,
         Tier = tier,
         `Proficency Level Cut`   = amo_lvl,
         Accuracy = accuracy,
         Consistency = consistency) %>% 
  arrange(content, Grade, Tier, `Proficency Level Cut`)
  
acc_con_tbls <- acc_con %>%   
  split(.$content) %>%
  map(~ungroup(.) %>% select(-content))
```

Results from the 2019-20 PASA test administration were analyzed using Rudner's
classification index (Rudner, 2005). Results closer to 1.0 indicate the
likelihood that a student was appropriately classified as proficient (Performance level of Proficient or Advanced) or not proficient (Performance Level of Novice or Emerging) and the
likelihood that the student would be classified in the same Performance level category given an additional test administration (Accuracy and Consistency, respectively). The
calculation utilizes item difficulty and (student ability) theta value
distributions, as well as related standard errors of measurement (SEM), to
generate probabilistic estimates based on one test administration. Complete
results for each content area and grade, generated from the `cacIRT` package in R,
are provided below. Please not that in cases in which there were no students in a proficiency level in a grade/tier, results cannot be reported.

Results indicate moderate to high levels of classification Accuracy (`r format(round(max(acc_con$Accuracy), 2), nsmall = 1)` to `r round(max(acc_con$Accuracy), 2)`) and Consistency (`r round(min(acc_con$Consistency), 2)` to `r format(round(max(acc_con$Consistency), 2), nsmall = 1)`).

```{r classification_accuracy_table, results="asis"}
walk2(acc_con_tbls,
      paste0(c("ELA", "Math", "Science"),
             " Accuracy & Consistency"),
      ~knitr::kable(.x, 
                    caption = .y,
                    digits = 2) %>%
        kable_styling(full_width = TRUE,
                      latex_options = "hold_position") %>%
        print())

```