---
title: "Chapter 8: Fairness and Accessibility"
author: "BRT"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(tidyverse)
require(janitor)
require(psych)
require(knitr)
require(kableExtra)
require(car)
require(here)

# This chapter requires the following data files in a folder called /score

# /datastore/2020PASA_ScalingAnalysis/enrollment/PASA Enrollment final 2020.csv

# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_ELA_Final_2020.csv
# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Math_Final_2020.csv
# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Science_Final_2020.csv

# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_ELA_final.csv
# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_Math_final.csv
# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_Science_final.csv

ela     <- read.csv(here::here("data", "score", "Scoring_ELA_Final_2020.csv"))
math    <- read.csv(here::here("data", "score", "Scoring_Math_Final_2020.csv"))
science <- read.csv(here::here("data", "score", "Scoring_Science_Final_2020.csv"))

ela2     <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_ELA_final.csv"))[,1:13]
math2    <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_Math_final.csv"))[,1:13]
science2 <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_Science_final.csv"))[.1:13]


ela3 <- merge(ela, ela2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade',
  'Tier'))

math3 <- merge(math, math2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade',
  'Tier'))

science3 <- merge(science, science2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade','Tier'))

enr <- read.csv(here::here("data", "score", "PASA Enrollment final 2020.csv"))

ela_enr <- merge(ela3, enr, by.x = 'Student_ID', by.y = 'PASA.Student.ID', all.x = TRUE, all.y = FALSE)

ela3$Gender <- factor(ela3$Gender, levels = c('F','M'), labels = c('Female','Male'))

math3$Gender <- factor(math3$Gender, levels = c('F','M'),labels = c('Female','Male'))

science3$Gender <- factor(science3$Gender, levels = c('F','M'),labels = c('Female','Male'))

ela3$ethnicity <- factor(ela3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian', 'Asian', 'Black or African American',
 'Hispanic','Multiracial','White'))

math3$ethnicity <- factor(math3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian','Asian','Black or African American', 'Hispanic','Multiracial','White'))

science3$ethnicity <- factor(science3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian', 'Asian', 'Black or African American', 'Hispanic','Multiracial','White'))

ela3$lep <- ifelse(ela3$lep == 0,'No', 'Yes')
math3$lep <- ifelse(math3$lep == 0,'No', 'Yes')
science3$lep <- ifelse(science3$lep == 0,'No', 'Yes')
ela3$fr_lunch <- ifelse(ela3$fr_lunch == 0,'NO', 'YES')
math3$fr_lunch <- ifelse(math3$fr_lunch == 0,'NO', 'YES')
```

# ELA

## Sex

Text summarizing the female-male difference (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}


  colnames(eff)[3]="Cohen's D (Female - Male)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)

```

### Appendix (SEX)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}
tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)
```


## Limited English Proficiency

Text summarizing the No - Yes difference for LEP (effect sizes can be used)...


```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (No - Yes)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)
```

### Appendix (LEP)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)
```

## Free Reduced Lunch

Text summarizing the No - Yes difference for FRL (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (No - Yes)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)
  
```

### Appendix (FRL)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)
```


## Disability

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$Disability), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(25:168, 1:24),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Disability', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 24) %>%
  pack_rows('Grade 4', 25, 48) %>%
  pack_rows('Grade 5', 49, 72) %>%
  pack_rows('Grade 6', 73, 96) %>%
  pack_rows('Grade 7', 97, 120) %>%
  pack_rows('Grade 8', 121, 144) %>%
  pack_rows('Grade 11', 145, 168)
```


## Ethnicity

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$ethnicity), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(13:84, 1:12),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Etnicity', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 12) %>%
  pack_rows('Grade 4', 13, 24) %>%
  pack_rows('Grade 5', 25, 36) %>%
  pack_rows('Grade 6', 37, 48) %>%
  pack_rows('Grade 7', 49, 60) %>%
  pack_rows('Grade 8', 61, 72) %>%
  pack_rows('Grade 11', 73, 84)
```


# Math


## Sex


Text summarizing the Female - Male difference (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$Gender),mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (Female - Male)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)
```

### Appendix (Sex)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$Gender),mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)

```


## Limited English Proficiency

Text summarizing the No - Yes difference for LEP (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')


eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (No - Yes)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)

```

### Appendix (LEP)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)
```

## Free Reduced Lunch

Text summarizing the No - Yes difference for FRL (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:14){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (No - Yes)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)
```


### Appendix (FRL)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 4) %>%
  pack_rows('Grade 4', 5, 8) %>%
  pack_rows('Grade 5', 9, 12) %>%
  pack_rows('Grade 6', 13, 16) %>%
  pack_rows('Grade 7', 17, 20) %>%
  pack_rows('Grade 8', 21, 24) %>%
  pack_rows('Grade 11', 25, 28)
```


## Ethnicity

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$ethnicity), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(13:84, 1:12),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Etnicity', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 12) %>%
  pack_rows('Grade 4', 13, 24) %>%
  pack_rows('Grade 5', 25, 36) %>%
  pack_rows('Grade 6', 37, 48) %>%
  pack_rows('Grade 7', 49, 60) %>%
  pack_rows('Grade 8', 61, 72) %>%
  pack_rows('Grade 11', 73, 84)
```

## Disability

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$Disability), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]

tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(25:168, 1:24),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Disability', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 3', 1, 24) %>%
  pack_rows('Grade 4', 25, 48) %>%
  pack_rows('Grade 5', 49, 72) %>%
  pack_rows('Grade 6', 73, 96) %>%
  pack_rows('Grade 7', 97, 120) %>%
  pack_rows('Grade 8', 121, 144) %>%
  pack_rows('Grade 11', 145, 168)
```


# Science

## Sex

Text summarizing the Female - Male difference (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score,
                  list(science3$Grade,
                       science3$Tier,
                       science3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(4,4,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:6){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (Female - Male)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)
  
```

### Appendix (Sex)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score,
                  list(science3$Grade,
                       science3$Tier,
                       science3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2, row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 4', 1, 4) %>%
  pack_rows('Grade 8', 5, 8) %>%
  pack_rows('Grade 11', 9, 12)
```

## Limited English Proficiency

Text summarizing the No - Yes difference for LEP (effect sizes can be used)...

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score, list(science3$Grade, science3$Tier, science3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(4,4,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2),
                  Cohen.d=NA)

for(i in 1:6){
  
  n1 = tab[2*i-1,]$N
  m1 = tab[2*i-1,]$Mean
  s1 = tab[2*i-1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1-m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3]="Cohen's D (No - Yes)"
  
  eff %>%
  kable(digits = 2, row.names = FALSE,align='c') %>%
  kable_styling(full_width = F)

```

### Appendix (LEP)

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score, list(science3$Grade, science3$Tier, science3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2,row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 4', 1, 4) %>%
  pack_rows('Grade 8', 5, 8) %>%
  pack_rows('Grade 11', 9, 12)
```

## Ethnicity

```{r, eval = TRUE, echo = FALSE, results = 'asis', warning = FALSE}

science3 <- rbind(
  science3,
  data.frame(Student_ID = NA,
             subject = NA,
             test_name = NA,
             Grade = 4,
             Tier = 1,
             Form.x = NA,
             Summed.Score = NA,
             No.Corr = NA,
             No.Inc = NA,
             No.Resp = NA,
             Not.Attempted = NA,
             PLC = NA,
             Scaled.Score = NA,
             Lower.Scale.Score = NA,
             Upper.Scale.Score = NA,
             StaffID = NA,
             UserID = NA,
             Form.y = NA,
             Disability = NA,
             Gender = NA,
             ethnicity = 'American Indian',
             lep = NA))

tab <- describeBy(science3$Scaled.Score, list(science3$Grade, science3$Tier, science3$ethnicity), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(13:nrow(tab), 1:12),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Etnicity', 'N', 'Mean', 'SD')

tab %>%
  kable(digits = 2,row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 4', 1, 12) %>%
  pack_rows('Grade 8', 13, 24) %>%
  pack_rows('Grade 11', 25, 36)
```


## Disability

```{r, eval = TRUE, echo = FALSE, results = 'asis', warning = FALSE}

tab <- describeBy(science3$Scaled.Score, list(science3$Grade, science3$Tier, science3$Disability), mat = T)
tab <- tab[,c('group1','group2','group3','n','mean','sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(25:nrow(tab), 1:24),]
tab <- tab[,-1]

colnames(tab) <- c('Tier','Disability','N','Mean','SD')

tab %>%
  kable(digits = 2,row.names = FALSE) %>%
  kable_styling(full_width = F) %>%
  pack_rows('Grade 4', 1, 24) %>%
  pack_rows('Grade 8', 25, 48) %>%
  pack_rows('Grade 11', 49, 72)
```

# School District Participation

The table below shows the participation in the 2019-2020 PASA ELA, Math, and Science content areas by school district.  

```{r, eval = TRUE, echo = FALSE, results = 'asis', warning = FALSE, message = FALSE}

enroll_raw <- rio::import(here("data", "score", "PASA Enrollment final 2020.csv")) %>%
  as_tibble() %>%
  clean_names() %>%
  select(tested_district_code, tested_school_code, student_id = pasa_student_id)

enroll <- enroll_raw %>%
  left_join(rio::import(here("data", "score", "districts and auns.xlsx")) %>%
  as_tibble() %>%
  clean_names() %>%
  rename(tested_district_code = service_provider_aun, district_name = district))

ela <- read_csv(here("data", "score", "Scoring_ELA_Final_2020.csv")) %>%
  clean_names()

math <- read_csv(here("data", "score", "Scoring_Math_Final_2020.csv")) %>%
  clean_names()

science <- read_csv(here("data", "score", "Scoring_Science_Final_2020.csv")) %>%
  clean_names()

district_dta <- ela %>%
  select(student_id, grade, tier, ela_score = scaled_score) %>%
  left_join(select(math, student_id, math_score = scaled_score)) %>%
  left_join(select(science, student_id, science_score = scaled_score)) %>%
  left_join(enroll)

district_table <- district_dta %>%
  mutate(ela_tot = sum(!is.na(ela_score)),
         math_tot = sum(!is.na(math_score)),
         science_tot = sum(!is.na(science_score))) %>%
  group_by(district_name) %>%
  summarise(ela_count = sum(!is.na(ela_score)),
            ela_pct = ela_count/unique(ela_tot)*100,
            math_count = sum(!is.na(math_score)),
            math_pct = math_count/unique(math_tot)*100,
            science_count = sum(!is.na(science_score)),
            science_pct = science_count/unique(science_tot)*100,)

district_table %>%
  janitor::adorn_totals(where = "row") %>%
  kable(., 
        digits = 2,
        #booktabs = TRUE,
        #longtable = TRUE, 
        col.names = c("School District", rep(c("Count", "Percent"), 3)),
        caption = "School District Participation in 2019-2020 PASA Across Content Area") %>%
  add_header_above(c(" " = 1, "ELA" = 2, "Math" = 2, "Science" = 2)) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```