---
title: Chapter 9
author: "BRT"
date: "`r format(Sys.Date(), '%m-%d-%Y')`"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
output: 
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: true
---

# Fairness and Accessibility

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      results = "asis",
                      message = FALSE)

require(tidyverse)
require(janitor)
require(psych)
require(knitr)
require(kableExtra)
require(car)
require(here)
library(gt)

# This chapter requires the following data files in a folder called /score

# /datastore/2020PASA_ScalingAnalysis/enrollment/PASA Enrollment final 2020.csv

# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_ELA_Final_2020.csv
# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Math_Final_2020.csv
# /datastore/2020PASA_ScalingAnalysis/scoring-files-2020/Scoring_Science_Final_2020.csv

# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_ELA_final.csv
# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_Math_final.csv
# /datastore/2020PASA_ScalingAnalysis/Assessment Scoring 2020/Assessment_Scoring_Data_Science_final.csv

ela     <- read.csv(here::here("data", "score", "Scoring_ELA_Final_2020.csv"))
math    <- read.csv(here::here("data", "score", "Scoring_Math_Final_2020.csv"))
science <- read.csv(here::here("data", "score", "Scoring_Science_Final_2020.csv"))

ela2     <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_ELA_final.csv"))[,1:13]
math2    <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_Math_final.csv"))[,1:13]
science2 <- read.csv(here::here("data", "score", "Assessment_Scoring_Data_Science_final.csv"))[.1:13]


ela3 <- merge(ela, ela2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade',
  'Tier'))

math3 <- merge(math, math2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade',
  'Tier'))

science3 <- merge(science, science2, by = c(
  'Student_ID',
  'subject',
  'test_name',
  'Grade','Tier'))

enr <- read.csv(here::here("data", "score", "PASA Enrollment final 2020.csv"))

ela_enr <- merge(ela3, enr, by.x = 'Student_ID', by.y = 'PASA.Student.ID', all.x = TRUE, all.y = FALSE)

ela3$Gender <- factor(ela3$Gender, levels = c('F','M'), labels = c('Female','Male'))

math3$Gender <- factor(math3$Gender, levels = c('F','M'),labels = c('Female','Male'))

science3$Gender <- factor(science3$Gender, levels = c('F','M'),labels = c('Female','Male'))

ela3$ethnicity <- factor(ela3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian', 'Asian', 'Black or African American',
 'Hispanic','Multiracial','White'))

math3$ethnicity <- factor(math3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian','Asian','Black or African American', 'Hispanic','Multiracial','White'))

science3$ethnicity <- factor(science3$ethnicity, levels = c('AI','AS','B','H','M','W'), labels = c('American Indian', 'Asian', 'Black or African American', 'Hispanic','Multiracial','White'))

ela3$lep <- ifelse(ela3$lep == 0,'No', 'Yes')
math3$lep <- ifelse(math3$lep == 0,'No', 'Yes')
science3$lep <- ifelse(science3$lep == 0,'No', 'Yes')
ela3$fr_lunch <- ifelse(ela3$fr_lunch == 0,'NO', 'YES')
math3$fr_lunch <- ifelse(math3$fr_lunch == 0,'NO', 'YES')
```

As noted in the Standards for Educational and Psychological Testing, Standard 3.3 specifies that “Those responsible for test development should include relevant subgroups in validity, reliability/precision, and other preliminary studies used when constructing the test” (p. 64). This argument is expanded to include reporting outcomes by relevant subgroups, noting however, that “subgroup differences do not in and of themselves indicate lack of fairness, but such differences should trigger follow-up studies, where feasible, to identify the potential causes of such differences” (p. 65). American Educational Research Association, American Psychological Association, National Council on Measurement in Education (2014). Standards for Educational and Psychological Testing. Washington, DC: Authors. This evaluation was completed by comparing the performance (by subject area, grade level, and Tier) for the following student groups: (a) sex, (b) race-ethnicity, (c) Limited English Proficiency, (d) disability, and (e) free-reduced price lunch. This evaluation involved comparing student counts and average differences (using the mean relative to the standard deviation). The tables are displayed in an appendix.  


# English Language Arts

## Gender

Few differences were reported by gender. Most of the student counts, averages, and standard deviations were similar for males and females. In Tier 2, slight differences (half a standard deviation) were present in the earlier (elementary) grades.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3] = "Cohen's D (Female - Male)"
```

```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```


```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (Female - Male)`))
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```


## Limited English Proficiency

With Limited English Proficiency (coded as No versus Yes) within Tiers and grade levels, few differences were apparent in averages and standard deviations. On the other hand, as expected, great differences were present in the number of students (count) in these two groups.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}
tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (No - Yes)"
```  
  
```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (No - Yes)`))
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```


## Economically Disadvantaged

For the analyses comparing students receiving free and reduced price lunch, slightly higher performances were present for those not receiving such support; this was true across grades as well as Tiers.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(ela3$Scaled.Score, list(ela3$Grade, ela3$Tier, ela3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

  colnames(eff)[3] = "Cohen's D (No - Yes)"
```

  
```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (No - Yes)`))
```

## Primary Disability

Like race-ethnicity (which had many different coded values), the analysis by disability resulted in great variation in counts, averages, and standard deviations. The two most present disabilities were Autism and Intellectual Disabilities, with Multiple Disabilities also quite prominent. This was true for both Tiers and across grade levels. Please see the appendix for descriptive statistics (counts, means, and standard deviations).

## Ethnicity

When comparing the performance of students from different racial and ethnic groups, across the grade levels and Tiers, considerable variation was present in all aspects of count (number of students), averages, and standard deviations. Please see the appendix for descriptive statistics (counts, means, and standard deviations).

# Mathematics

## Gender

We found more males participating in the PASA than females, sometimes double the number. This was true across grades and Tiers. Furthermore, performance differences were slightly greater for males than females, across grades and Tiers. Both males and females had similar variance (standard deviations).

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$Gender),mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (Female - Male)"
```


```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```


```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (Female - Male)`))
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

## Limited English Proficiency

Limited English Proficient students, though fewer than the number of students deemed proficient, performed as well to those proficient, in all grades and Tiers; the variance in these two groups was comparable.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')


eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (No - Yes)"
```

```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (No - Yes)`))
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```


## Economically Disadvantaged

Students receiving free and reduced price lunch performed comparably to students who did not receive such support. The count of students in both groups was much greater in Tier 1. Though fewer in number (by half), the averages and standard deviations were quite similar with those receiving support often a bit higher (by a half standard deviation).

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(math3$Scaled.Score, list(math3$Grade, math3$Tier, math3$fr_lunch), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:28, 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'FRL', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(3,3,4,4,5,5,6,6,7,7,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2,1,2,1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:14) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (No - Yes)"
```

  
```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (No - Yes)`))
```


## Ethnicity

The largest group of students were White, followed by Black/African American, and finally Hispanic students. This was true for all grades and Tiers. Wide variation existed in both the averages and standard deviations. Please see the appendix for descriptive statistics (counts, means, and standard deviations).

## Primary Disability

The largest group of students were with Intellectual Disabilities or Autism, followed by Multiple Disabilities. Great variation was present in averages and standard divisions. Many disabilities had very few students participating in the PASA. Please see the appendix for descriptive statistics (counts, means, and standard deviations).

# Science

## Gender

In most grades and Tiers, far more males than females participated in the Science PASA and for both groups, more students took Tier 2 than Tier 1. Their performance (averages and standard deviations) were comparable.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score,
                  list(science3$Grade,
                       science3$Tier,
                       science3$Gender), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'Sex', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(4,4,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:6) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (Female - Male)"
```

```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (Female - Male)`))
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

## Limited English Proficiency

The number of students with Limited English Proficiency was fairly small and much less than those with English proficiency. Both groups performed comparably (in averages and standard deviations) across the grades and Tiers.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}

tab <- describeBy(science3$Scaled.Score, list(science3$Grade, science3$Tier, science3$lep), mat = T)
tab <- tab[,c('group1', 'group2', 'group3', 'n', 'mean', 'sd')]
tab <- tab[with(tab, order(group1, group2)),]
tab <- tab[c(5:nrow(tab), 1:4),]
tab <- tab[,-1]

colnames(tab) <- c('Tier', 'LEP', 'N', 'Mean', 'SD')

eff <- data.frame(Grade = c(4,4,8,8,11,11), 
                  Tier  = c(1,2,1,2,1,2),
                  Cohen.d = NA)

for (i in 1:6) {
  
  n1 = tab[2*i - 1,]$N
  m1 = tab[2*i - 1,]$Mean
  s1 = tab[2*i - 1,]$SD
  
  n2 = tab[2*i,]$N
  m2 = tab[2*i,]$Mean
  s2 = tab[2*i,]$SD
  
  eff[i,3] = (m1 - m2)/sqrt((s1^2 + s2^2)/2)
}

colnames(eff)[3] = "Cohen's D (No - Yes)"
```

```{r eval = knitr::is_latex_output()}
eff %>%
  kable("latex", booktabs = TRUE, digits = 2, row.names = FALSE, align = 'c') %>%
  kable_styling(full_width = FALSE)
```

```{r eval = knitr::is_html_output()}
gt(eff) %>% 
  fmt_number(vars(`Cohen's D (No - Yes)`))
```


## Ethnicity

The count was greatest with White, then Black-African American, and Hispanic students. Values (averages and standard deviations) were slightly varied across the grades and Tiers but not as much as in ELA and Mathematics. Please see the appendix for descriptive statistics (counts, means, and standard deviations).

## Primary Disability

As in ELA and Mathematics, the count of students was greatest with Autism and Intellectual Disabilities. Performance was somewhat varied (in averaged and standard deviations) across the grades and Tiers. Some disability areas were represented with only a few students, making it difficult to generalize. Please see the appendix for descriptive statistics (counts, means, and standard deviations).


# Appendix 

Note - cell sizes less than 5 not reported.

```{r, eval = TRUE, echo = FALSE, results = 'asis'}
appendix <- bind_rows(ela3, math3, science3) %>% 
  as_tibble() %>% 
  select(subject, Grade, Tier, Gender, lep, fr_lunch, Disability,
         ethnicity, Scaled.Score) %>% 
  pivot_longer(c(Gender, lep, fr_lunch, Disability, ethnicity),
               names_to = "demo", 
               values_to = "Group") %>% 
  group_by(subject, Grade, Tier, demo, Group) %>% 
  summarize(n = n(),
            mean = mean(Scaled.Score, na.rm = TRUE),
            sd = sd(Scaled.Score, na.rm = TRUE)) %>% 
  mutate(mean = ifelse(n < 5, NA_real_, mean),
         sd = ifelse(n < 5, NA_real_, sd)) %>% 
  ungroup()
```

## English/Language Arts

```{r}
ela_appendix <- appendix %>% 
  filter(subject == "ELA") %>% 
  select(-subject) %>% 
  group_by(Grade, Tier, demo) %>% 
  nest()

ela_gender <- filter(ela_appendix, demo == "Gender")
ela_lep <- filter(ela_appendix, demo == "lep")
ela_frl <- filter(ela_appendix, demo == "fr_lunch")
ela_dis <- filter(ela_appendix, demo == "Disability")
ela_eth <- filter(ela_appendix, demo == "ethnicity")
```

### Gender

```{r eval = knitr::is_html_output()}
pwalk(list(ela_gender$Grade, 
           ela_gender$Tier, 
           ela_gender$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(ela_gender$Grade, 
           ela_gender$Tier, 
           ela_gender$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Limited English Proficient

```{r eval = knitr::is_html_output()}
pwalk(list(ela_lep$Grade, 
           ela_lep$Tier, 
           ela_lep$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(ela_lep$Grade, 
           ela_lep$Tier, 
           ela_lep$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Economically Disadvantaged

```{r eval = knitr::is_html_output()}
pwalk(list(ela_frl$Grade, 
           ela_frl$Tier, 
           ela_frl$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(ela_frl$Grade, 
           ela_frl$Tier, 
           ela_frl$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

### Primary Disability

```{r eval = knitr::is_html_output()}
pwalk(list(ela_dis$Grade, 
           ela_dis$Tier, 
           ela_dis$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(ela_dis$Grade, 
           ela_dis$Tier, 
           ela_dis$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

### Race/Ethnicity

```{r eval = knitr::is_html_output()}
pwalk(list(ela_eth$Grade, 
           ela_eth$Tier, 
           ela_eth$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(ela_eth$Grade, 
           ela_eth$Tier, 
           ela_eth$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

## Mathematics

```{r}
math_appendix <- appendix %>% 
  filter(subject == "Math") %>% 
  select(-subject) %>% 
  group_by(Grade, Tier, demo) %>% 
  nest()

math_gender <- filter(math_appendix, demo == "Gender")
math_lep <- filter(math_appendix, demo == "lep")
math_frl <- filter(math_appendix, demo == "fr_lunch")
math_dis <- filter(math_appendix, demo == "Disability")
math_eth <- filter(math_appendix, demo == "ethnicity")
```

### Gender

```{r eval = knitr::is_html_output()}
pwalk(list(math_gender$Grade, 
           math_gender$Tier, 
           math_gender$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(math_gender$Grade, 
           math_gender$Tier, 
           math_gender$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Limited English Proficient

```{r eval = knitr::is_html_output()}
pwalk(list(math_lep$Grade, 
           math_lep$Tier, 
           math_lep$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(math_lep$Grade, 
           math_lep$Tier, 
           math_lep$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Economically Disadvantaged

```{r eval = knitr::is_html_output()}
pwalk(list(math_frl$Grade, 
           math_frl$Tier, 
           math_frl$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(math_frl$Grade, 
           math_frl$Tier, 
           math_frl$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

### Primary Disability

```{r eval = knitr::is_html_output()}
pwalk(list(math_dis$Grade, 
           math_dis$Tier, 
           math_dis$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(math_dis$Grade, 
           math_dis$Tier, 
           math_dis$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

### Race/Ethnicity

```{r eval = knitr::is_html_output()}
pwalk(list(math_eth$Grade, 
           math_eth$Tier, 
           math_eth$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(math_eth$Grade, 
           math_eth$Tier, 
           math_eth$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

## Science

```{r}
sci_appendix <- appendix %>% 
  filter(subject == "Science") %>% 
  select(-subject) %>% 
  group_by(Grade, Tier, demo) %>% 
  nest()

sci_gender <- filter(sci_appendix, demo == "Gender")
sci_lep <- filter(sci_appendix, demo == "lep")
sci_frl <- filter(sci_appendix, demo == "fr_lunch")
sci_dis <- filter(sci_appendix, demo == "Disability")
sci_eth <- filter(sci_appendix, demo == "ethnicity")
```

### Gender

```{r eval = knitr::is_html_output()}
pwalk(list(sci_gender$Grade, 
           sci_gender$Tier, 
           sci_gender$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(sci_gender$Grade, 
           sci_gender$Tier, 
           sci_gender$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Limited English Proficient

```{r eval = knitr::is_html_output()}
pwalk(list(sci_lep$Grade, 
           sci_lep$Tier, 
           sci_lep$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(sci_lep$Grade, 
           sci_lep$Tier, 
           sci_lep$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r eval = knitr::is_html_output(), results = "asis"}
cat("<br/>")
```

### Economically Disadvantaged

```{r eval = knitr::is_html_output()}
pwalk(list(sci_frl$Grade, 
           sci_frl$Tier, 
           sci_frl$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(sci_frl$Grade, 
           sci_frl$Tier, 
           sci_frl$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```

### Primary Disability

```{r eval = knitr::is_html_output()}
pwalk(list(sci_dis$Grade, 
           sci_dis$Tier, 
           sci_dis$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(sci_dis$Grade, 
           sci_dis$Tier, 
           sci_dis$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


### Race/Ethnicity

```{r eval = knitr::is_html_output()}
pwalk(list(sci_eth$Grade, 
           sci_eth$Tier, 
           sci_eth$data), ~{
  cat("<br/>")
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  
  gt(..3) %>% 
    fmt_number(vars(mean, sd)) %>% 
    print()
})
```

```{r eval = knitr::is_latex_output()}
pwalk(list(sci_eth$Grade, 
           sci_eth$Tier, 
           sci_eth$data), ~{
  cat("\n",
      paste("#### Grade", ..1, "Tier", ..2, "\n"),
      sep = "\n")
  cat("\\leavevmode")
  kable(..3, "latex", booktabs = TRUE, digits = 2) %>% 
    kable_styling(full_width = FALSE) %>% 
    print()
})
```


```{r, include=knitr::is_html_output(), results = "asis"}
cat('
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
'
)
```